---
title: "A qualitative study of mechanisms underlying effects of a parenting intervention in rural Liberia"
output:
  html_document:
    toc_float: true
bibliography: parenting.bib
csl: vancouver.csl
---

# Metadata

```{r, include=FALSE}
  library(magrittr)
  library(knitr)
  library(kableExtra)
  library(tidyverse)
```

```{r, echo=FALSE}
  tbl <- data.frame(Element=c("Paper Citation", 
                              "Journal Copy", 
                              "Pre-Print"),
                    Details=c("Giusto, A., Friis, E., Sim, A. L., Chase, R. M., Zayzay, J. O., Green, E., & Puffer, E. (2017). A qualitative study of mechanisms underlying effects of a parenting intervention in rural Liberia. The European Journal of Development Research, 29(5), 964-982.",
                              "[The European Journal of Development Research](https://link.springer.com/article/10.1057/s41287-017-0101-8)",
                              "[Research Gate](https://www.researchgate.net/profile/Ali_Giusto/publication/319920721_A_Qualitative_Study_of_Mechanisms_Underlying_Effects_of_a_Parenting_Intervention_in_Rural_Liberia/links/5a123f180f7e9bd1b2c10257/A-Qualitative-Study-of-Mechanisms-Underlying-Effects-of-a-Parenting-Intervention-in-Rural-Liberia.pdf?_sg%5B0%5D=5y1r6K_iLI0_bhFJgbmsS2HzdgfXVstFcDVQu0ivBtnJmXLheeSdyiE7xKKaHbtCJUX_93ysHdAB1pv80U-Qng.DniHxczU_LaSUDjBXVRmWwm4x38p08sF4VuHqc5a4zSjn3K2C9pfu9YmPNK76wQxi2k4SYWeb_8bFPkVDc5A2g&_sg%5B1%5D=RaD6EeFRtD2mSqf5lTrTi5A0A9Ep68q4sgagZffPwRBlNxfIgiDYBvMYBey-4UUEN8os5HM9hDCwhesXwstz5Ygd1jAOcZw8oEyfRKIA3pXz.DniHxczU_LaSUDjBXVRmWwm4x38p08sF4VuHqc5a4zSjn3K2C9pfu9YmPNK76wQxi2k4SYWeb_8bFPkVDc5A2g&_iepl=)"))

tbl %>%
  kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>%
  column_spec(1, width = "10em")
```

# Summary

Parenting interventions can reduce child maltreatment and improve child outcomes in high-risk settings, but little is known about mechanisms underlying effects. This study presents qualitative findings on mechanisms of change from a randomized trial of a parenting intervention in Liberia. Participants (N = 30) completed semi-structured interviews, and thematic content analysis was conducted from transcripts. Results suggest that learning about effects of violence on child development and discussing the value of empathy for children strengthened caregivers’ sense of identity as nurturers and protectors. This in turn drove efforts to decrease harsh discipline. As a result, children expressed less fear, increasing opportunities for positive interactions; shared enjoyment maintained reduced harsh treatment. Caregivers also described recognizing that physical punishment was often ineffective and using new non-violent discipline skills alongside emotion regulation skills to facilitate behavior change. Participants also described reduced couples conflict and more peaceful home environments associated with increased self-identification as role models.