---
title: "The impact of the 2014 Ebola virus disease outbreak in Liberia on parent preferences for harsh discipline practices: a quasi-experimental, pre-post design"
output:
  html_document:
    toc_float: true
bibliography: parenting.bib
csl: vancouver.csl
---

# Metadata

```{r, include=FALSE}
  library(magrittr)
  library(knitr)
  library(kableExtra)
  library(tidyverse)
```

```{r, echo=FALSE}
  tbl <- data.frame(Element=c("Paper Citation", 
                              "Journal Copy", 
                              "Pre-Print", 
                              "Data Citation"),
                    Details=c("Green, E., Chase, R. M., Zayzay, J., Finnegan, A., & Puffer, E. S. (2018). The impact of the 2014 Ebola virus disease outbreak in Liberia on parent preferences for harsh discipline practices: a quasi-experimental, pre-post design. Global Mental Health, 5.",
                              "[Global Mental Health](https://www.cambridge.org/core/journals/global-mental-health/article/impact-of-the-2014-ebola-virus-disease-outbreak-in-liberia-on-parent-preferences-for-harsh-discipline-practices-a-quasiexperimental-prepost-design/80ECA35B6338DE7EF26DD60FE25702E6)",
                              "[PMC](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5797936/)",
                              "Green, E.P. & Puffer, E.P. (2017). PMD data repository. [![DOI](https://zenodo.org/badge/116026154.svg)](https://zenodo.org/badge/latestdoi/116026154)"))

tbl %>%
  kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>%
  column_spec(1, width = "10em")
```

# Summary

This paper uses data from a cohort of parents and guardians of young children living in Monrovia, Liberia collected before and after the 2014 outbreak of Ebola virus disease (EVD) to estimate the impact of EVD exposure on implicit preferences for harsh discipline. We hypothesized that parents exposed to EVD-related sickness or death would exhibit a stronger preference for harsh discipline practices compared with non-exposed parents.

The data for this analysis come from two survey rounds conducted in Liberia as part of an intervention trial of a behavioral parenting skills intervention. Following a baseline assessment of 201 enrolled parents in July 2014, all program and study activities were halted due to the outbreak of EVD. Following the EVD crisis, we conducted a tracking survey with parents who completed the baseline survey 12 months prior. In both rounds, we presented parents with 12 digital comic strips of a child misbehaving and asked them to indicate how they would react if they were the parent in the stories.

Parents from households with reported EVD sickness or death became more ‘harsh’ (Glass's delta = 1.41) in their hypothetical decision-making compared with non-exposed parents, t (167)=−2.3, p  <  0.05. Parents from households that experienced EVD-related sickness or death not only reported significantly more household conflict and anxiety, but also reported that their child exhibited fewer difficulties.

Results support the need for family-based interventions, including strategies to help parents learn alternatives to harsh punishment.